---
title: "Assignment04"
author: "Farah Kaddah"
date: "10/3/2021"
output: html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, include=FALSE}
library (tidyverse)
library(ggplot2)
library(dplyr)
library(readr)
library(dplyr)
library(sf)
library(jsonlite)
library(httr)
library(viridis)
library(hrbrthemes)
```

## Mapping Conflict in Syria by event type

```{r Loading, echo=TRUE, results = 'hide'}

conflicts <- read.csv("Data/ACLED_2018-09-10-2021-09-17-Eastern_Africa-Middle_Africa-Middle_East-Southern_Africa-Western_Africa.csv") %>%
  st_as_sf(coords = c("longitude", "latitude")) %>%
  filter (country== "Syria")

maps <-  st_read ("Data/World_Countries") %>%
          filter (COUNTRY == "Syria")

```


```{r map, echo=TRUE}

conflicts <- st_set_crs(conflicts, st_crs(maps))
Syria <- ggplot() +
  geom_sf(
    data = maps, size = 0.1, color = "black", fill = "NA") +
  geom_sf (data = filter (conflicts, year == 2021), aes(color= event_type), alpha=0.3,  size=0.5) +
  labs(
    title = "Syria",
    subtitle = "Conflict Distribution by event type in Syria (2021)",
    caption = "Source: ACLED Data
    link: https://acleddata.com/data-export-tool/") +
     theme_void() +
       scale_colour_discrete("Event Types") 

# making legend items bigger
Syria +
  guides(color = guide_legend(override.aes = list(size=3, alpha=1)))
      

```

The visual above shows the distribution of different types of conflicts in Syria in 2021 until September. Violence against civilians and explosions/remote violence seems to be the dominant conflict events in Syria. Conflict however is not evenly distributed across the country - with significant clustering around the north around and the south west. Given the dominant types of conflict in Syria, mapping where most of it occurs is expected to help humanitarian aid planning as well as international intervention. 

## Stacked Bar Chart: Regional Conflict Data by Event Type
### Middle East and Africa

```{r ACLED, echo=TRUE}

conflict_bar <- read.csv("Data/ACLED_2018-09-10-2021-09-17-Eastern_Africa-Middle_Africa-Middle_East-Southern_Africa-Western_Africa.csv")

```

```{r cleaning, echo=TRUE, results='hide'}

clean_conflict <- select(conflict_bar, event_type, region, year) %>%
                  filter(year== 2021) %>% 
                  group_by(region)  %>%
                  count(n=event_type)  
                 
```

```{r type, echo=TRUE}

clean_conflict %>%
  ggplot(aes(fill= n, y=nn, x=region)) + 
  geom_bar(position="stack", stat="identity")+
  xlab("Region") +
  ylab("Number of Events") +
  scale_colour_discrete("Event Types")+
  labs(title = "Conflict in Middle East and Africa", subtitle = "By conflict type data, 2021", caption = "Source: ACLED data
       link: https://acleddata.com/data-export-tool/") +
  guides(fill=guide_legend(title="Event Type")) +
  theme_minimal() +
  theme(
    plot.title = element_text(color = "#0099f9", size = 20),
    plot.subtitle = element_text(face = "bold"),
    plot.caption = element_text(face = "italic")
  )

```

The stacked bar char shows a comparison between the Middle East as well as Africa in 2021 (until September). It shows the number of conflict events occurring in each region with further categorization in to types: battles, explosions/remote Violence, protests, riots, strategic developments, and  violence against civilians. The Middle East has had the highest number of conflict events in 2021, most likely driven by Syria, Palestine and Yemen. Protests and explosions/remote violence are the most reported events in the Middle East which is not the case in other African regions. The literature suggests poverty to be a main driver of conflict in developing countries. Although the Middle East and African regions are both suffering from developmental issues, conflict manifest themselves differently. This suggests looking the structural differences of these developmental problems across the regions, or possibly other reasons, to identify these differences in the distribution of conflict events. 


# Correlation between HDI & Corruption
```{r qog, echo=TRUE}

qog_std_cs_jan21 <- read_csv("Data/qog_std_cs_jan21.csv")

#Scatter plot for HDI and corruption
qof_filtered <- select (qog_std_cs_jan21, undp_hdi, cname, wef_cor,vdem_corr)

```

```{r plot, echo=TRUE}
qof_filtered %>%
  ggplot(aes(x=undp_hdi, y=wef_cor)) + 
  geom_point(color= "blue", size= 2, alpha = 0.5) +
  geom_smooth(method=lm, se=TRUE, fullrange=TRUE, level=0.95, color ="black") +
  scale_x_continuous("Human Development Index (0-1)", limits=c(0.2,1))+
  scale_y_continuous("Incidence of corruption (0 -100)", limits=c(0,100)) +
  labs(title = "Global Relationship between Corruption and HDI 2020", subtitle = "Countries with higher HDI tend to have lower corruption", caption = " Source: Quality of Government Dataset
       https://www.gu.se/en/quality-government/qog-data/data-downloads/standard-dataset" ) +
  theme_minimal()

```

The scatter plot above shows a relatively strong positive correlation between human development and "clean" countries globally. The Corruption Perceptions Index aggregates data from a number of different sources that provide perceptions of business people and country experts of the level on corruption in the public sector. The scale ranges from 0 [highly corrupt] to 100 [very clean]. It is difficult to say from the scatter plot alone without more tests which variable drives the other. Corruption on one had halts development because it creates a lot of dead weight loss in the economy and erodes trust in the system. One could still argue that countries with high HDI have "cleaner" environments because of their strong development reflected in an effective institutional set up and transparency. 


## HDI Histogram

```{r histogram, echo=TRUE, message=FALSE, warning=FALSE}
qof_filtered %>%
  ggplot (aes(x=undp_hdi)) +
    geom_histogram (binwidth=0.02, fill="#69b3a2", color="#e9ecef", alpha=0.9)+
   labs(title = "Global Human Development Distribution 2020", subtitle = "Skewed left distribution", 
        caption = " Source: Quality of Government Dataset
        link: https://www.gu.se/en/quality-government/qog-data/data-downloads/standard-dataset") +
    xlab("HDI")+
    ylab("Number of Countries") +
    scale_x_continuous(limits=c(0.3,1)) +
    theme_minimal()+
    theme(
    plot.title = element_text(size=15)
    )
```

The histogram above shows the distribution of countries across the human development index. The histogram is skewed to the left with a modal HDI around 0.7. No countries lie below an HDI of 0.3, while most countries seem to be concentrated above the 0.6 HDI mark.There is a stark difference between the top performing countries (HDI around 0.9)  and the lowest performing countries (HDI around 0.4). While HDI has been on the rise for the past 50 years, this graph shows that the world is still suffering from extra inequality between countries. Linking to the graphs on conflict above, it would be interesting as an additional task later to see correlations between inequality and conflict. 

